---
title: "Cost-Benefit Analysis"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = 'figures/')
```

- use values from confusion matrix for one representative model on a test set. or get average precision? choose a particular sensitivity level and see the range in precision?
  - Number needed to screen (NNS) - the number of alerted patients the models must flag to identify 1 true positive. 
  - Number needed to treat (NNT) - the number of true positive patients one must treat for 1 patient to benefit from the treatment.
  - Number needed to benefit (NNB = NNS x NNT) - how many patients must be screened for 1 patient to benefit.
- assumptions
  - cost of a non-severe case
  - cost of a severe case
    - days in icu, colectomy
  - cost of possible treatments:
    - abx: metronizadole, vancomycin, fidaxomycin
    - fmt, bezlotoxumab (monoclonal antibodies) 


## evaluating ml models

### [number needed to benefit](https://academic.oup.com/jamia/article-abstract/26/12/1655/5516459)

- Number needed to screen (NNS) - the number of alerted patients the models must flag to identify 1 true positive. 
- Number needed to treat (NNT) - the number of true positive patients one must treat for 1 patient to benefit from the treatment.
- Number needed to benefit (NNB = NNS x NNT) - how many patients must be screened for 1 patient to benefit.

> In the simplest terms, prediction can be distilled into an NNS and action into a number needed to treat. When contextualized within this framework, the product of NNS and number needed to treat results in a number needed to benefit. The table outlines key variables in this framework that will alter the estimated number needed to benefit across different modeling and implementation scenarios.

### [decision curve analysis](https://journals.sagepub.com/doi/10.1177/0272989X06295361)

QALYs are prone to systemic biases. 
We can sidestep using them by doing decision curve analysis instead.
Weight the relative harms of FP and FN results.

- incidence of severe CDI
- net benefit = $\frac{TP}{n} - \frac{FP}{n} \times \frac{p_t}{1-p_t}$ (Peirce 1884)
- plot net benefit vs probability threshold ($p_t$).
- harm of treating with fidaxomicin instead of vancomycin: more expensive. by how much?
- assume all patients severe vs no patients severe.

```{r decision-curve}
library(here)
library(schtools)
library(tidyverse)
thresh_dat <- read_csv(here('results','thresholds_results_aggregated.csv'))

max_thresholds <- thresh_dat %>% 
  filter(strategy == 'model') %>% 
  group_by(dataset, outcome, strategy, decision_threshold)  %>% 
  summarize(mean_nb = mean(net_benefit)) %>% 
  filter(mean_nb < 0) %>% 
  slice_min(mean_nb) %>% 
  ungroup() %>% 
  mutate(max_thresh = decision_threshold) %>% 
  select(dataset,outcome,max_thresh)

thresh_dat %>% 
  left_join(max_thresholds, by = join_by(dataset, outcome)) %>% 
  filter(decision_threshold <= max_thresh) %>% 
  group_by(dataset, outcome, strategy, decision_threshold) %>% 
  summarize(mean_nb = mean(net_benefit)) %>% 
  ggplot(aes(decision_threshold, mean_nb, 
             color = strategy, linetype = strategy)) +
  geom_line(alpha = 0.6) +
  facet_grid(dataset ~ outcome, scales = 'free') +
  theme_sovacool()
```


## treatment options & costs

- Vince: consider bezlotoxumab since it targets the toxin, mouse models show it reduces organ damage.
- Krishna: bezlotoxumab clinical trials didn't find signal for limiting severity. however fidaxomicin is superior to vancomycin for cure and time to resolution of diarrhea ([IDSA 2021 update](https://doi.org/10.1093/cid/ciab549)).
  - Fidaxomicin [NNT=10](https://thennt.com/nnt/oral-fidaxomicin-versus-vancomycin-clostridioides-difficile-infection/) for global cure (no recurrence occured).

NNT for Fidaxomicin, FMT, and/or Bezlotoxumab. Current standard is Vancomycin
because it's cheaper than Fidaxomicin, even though IDSA recommends Fidaxomicin.

rough estimate of costs.
current: everyone gets vancomycin.
potential: patients flagged as severe get fidaxomicin. based on NNB, estimate
how much money saved in averting severe outcomes.

### [Gupta _et al._ 2021 - Economic Burden of CDI](https://journals.sagepub.com/doi/10.1177/17562848211018654)

- average cost of CDI case in the US: 
  - \$8k to \$30k (Nanwa _et al_)
  - \$21k (Zhang _et al_)
  - likely underestimates of true attributable costs.
- treatment   
  - IDSA recommends either vancomycin or fidaxomycin for 10 days
  - Metronizadole out of favor, not efficacious
  - Recommend FMT after multiple recurrences
  - monoclonal Ab now an fda-approved treatment

#### treatment costs

> More recently, Rajasingham et al. calculated the costs of the currently
> available therapies for CDI.
> The cost of oral metronidazole (10-day course) ranged from US$4.38 to
> US$13.14, intravenous metronidazole (14-day course) from US$19.56 to $58.68,
> vancomycin (10-day course) from US$7.04 to US$21.12, rifaximin (20-day course)
> from US$44.16 to US$132.48, and fidaxomicin (10day course), being the most
> expensive option, ranged from US$883.60 to US$2650.80. It is difficult to
> predict the exact cost of FMT due to the multiple variables involved, including
> source of stool and route of administration. In general, one course of FMT is
> estimated to cost between US$500 and US$2000.


